---
title: "Heart Attack Risk Prediction. Final Project"
author: "Giselle Rahimi, Pilar Gore, WonJune Lee, Jiseon Kim, Jiwon Jung, Yoon-hye Cho, Sukyoung
  Yoon"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(modelr)
library(broom)
library(ggmosaic)
library(boot)
library(infer)
library(effsize)
library(plotly)
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(geosphere))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(plotrix))
suppressPackageStartupMessages(library(ggpubr))
library(rsample)

library(tidymodels)
load("bootstrap_cohens_d.RData")


untidied_dataset <- read_csv("heart_attack_prediction_dataset.csv")

cost <- function(obs, pred) {
  outcome <- pred > 0.5
  return(1 - (sum(obs == outcome) / length(obs)))
}

```

#Tidying (Yoonhye)
#Removed unwanted columns

```{r}
untidied_dataset <- untidied_dataset %>%
select(Age,
       Sex,
       Cholesterol,
       `Blood Pressure`,
       `Heart Rate`,
       `Family History`,
       `Exercise Hours Per Week`,
       `Stress Level`,
       `Sleep Hours Per Day`,
       `Heart Attack Risk`,
       `Triglycerides`,
       `Physical Activity Days Per Week`,
       `Sedentary Hours Per Day`,
       `Diet`)
```

#Renamed columns:

```{r}
untidied_dataset <- untidied_dataset %>%
  rename(
    heart_attack_risk = "Heart Attack Risk",
    blood_pressure = "Blood Pressure",
    heart_rate = "Heart Rate",
    family_history = "Family History",
    exercise_hrs_week = "Exercise Hours Per Week",
    stress_level = "Stress Level",
    sedentary_hrs_day = "Sedentary Hours Per Day",
    physical_days_week = "Physical Activity Days Per Week",
    sleep_hrs = "Sleep Hours Per Day",
    age = "Age",
    sex = "Sex",
    cholesterol = "Cholesterol",
    diet = "Diet",
    triglycerides = "Triglycerides"
  )

```

#Split original dataset into "no family history" and "family history":

```{r}
no_family_history_dataset <- 
  subset(untidied_dataset,
         family_history == 0)
```

```{r}
family_history_dataset <- 
  subset(
    untidied_dataset,
    family_history == 1)
```

#Separated the blood_pressure column into systolic and diastolic:

```{r}
no_family_history_dataset <- no_family_history_dataset %>%
  separate(
    col = blood_pressure,
    into = combine("systolic",
                   "diastolic"),
    sep = "/",
    convert = FALSE
  )

```
#EDA (Jiwon Jung)
#boxplot for the exercise hours explanatory variable:

```{r}
no_family_history_dataset %>%
ggplot() +
  geom_boxplot(
    mapping = aes(
      x = 
        factor(heart_attack_risk),
      y = exercise_hrs_week))+
  labs(
    title = "Boxplot of Heart Attack Risk and 
    Exercise Hours per Week",
    x = "Heart Attack Risk?",
    y = "Exercise Hours per Week")

```
#summary statistics for exercise hours:

```{r}
no_family_history_dataset %>%
  group_by(
    factor(heart_attack_risk)) %>%
    summarize(
    median = median(exercise_hrs_week),
    mean = mean(exercise_hrs_week),
    max = max(exercise_hrs_week),
    min = min(exercise_hrs_week)
  )
```
#Boxplot for sleep hours explanatory variable:

```{r}
no_family_history_dataset %>%
  ggplot() +
  geom_boxplot(
    mapping = aes(
      x = factor(heart_attack_risk),
      y = sleep_hrs)) +
  labs(
    title = "Boxplot of Heart Attack Risk 
    and Sleep Hours per Day", 
    x = "Heart Attack Risk?",
    y = "Sleep Hours")
```
#Summary statistics of sleep hours:

```{r}
no_family_history_dataset %>%
  group_by(
    factor(heart_attack_risk)) %>%
    summarize(
    median = median(sleep_hrs),
    mean = mean(sleep_hrs),
    max = max(sleep_hrs),
    min = min(sleep_hrs)
  )
```
#Boxplot of stress level explanatory variable:

```{r}
no_family_history_dataset %>%
  ggplot() +
  geom_boxplot(
    mapping = aes(
      x = factor(heart_attack_risk),
      y = stress_level)) +
  labs(
    title = "Boxplot of Heart Attack
    Risk and Stress Level",
    x = "Heart Attack Risk",
    y = "stress level")
```
#Summary statistics of stress level explanatory variable:

```{r}
no_family_history_dataset %>%
  group_by(
    factor(heart_attack_risk)) %>%
    summarize(
    median = median(stress_level),
    mean = mean(stress_level),
    max = max(stress_level),
    min = min(stress_level)
  )
```

#Splitting the stress levels into "not stressed" and "stressed":

```{r}
no_family_history_dataset <- no_family_history_dataset %>%
  mutate(
    count = 
      ifelse(
        stress_level >= 1 
        & !(stress_level >=6) 
        & heart_attack_risk == 1,
        "no stress & risk",
      ifelse(
        stress_level >= 1
        & !(stress_level >=6)
        & heart_attack_risk == 0,
        "no stress & no risk",
      ifelse(
        stress_level >= 6 
        & heart_attack_risk == 1,
        "stress & risk",
      ifelse(
        stress_level >= 6 
        & heart_attack_risk == 0,
        "stress and no risk", NA))))
  )
```

#Counted how many people were stressed and at a risk of heart attacks:

```{r}
no_family_history_dataset %>%
  group_by(factor(count)) %>%
  summarize(n = n())
```

#Bar blot showing distribution of sleep Hours:

```{r}
no_family_history_dataset$sleep_hrs <- as.numeric(no_family_history_dataset$sleep_hrs)

```

```{r}
no_family_history_dataset %>%
  ggplot()+
  geom_histogram(
    mapping = aes(x = sleep_hrs,
                  y = ..density..),
    bins = 7,
    fill = "light blue",
    color = "black")+
  labs(
    title = "Distribution of Sleep Hours",
    x = "Sleep Hours")+
  facet_wrap(
    ~heart_attack_risk,
    scales = "free_x")
```

#PMF plot showing distribution of Exercise Hours per week:


```{r}
no_family_history_dataset %>%
  ggplot() +
  geom_histogram(
    mapping = aes(
      x = exercise_hrs_week,
      y = ..density..),
    bins = 10, 
    fill = "light blue",
    color = "black")+
  labs(
    title = 
      "PMF distribution of Exercise hours per week")+
  facet_wrap(
    ~heart_attack_risk,
    scales = "free_x")
```
#Histogram showing distribution of stress level:

```{r}
no_family_history_dataset$stress_level <- as.numeric(no_family_history_dataset$stress_level)

```

```{r}
no_family_history_dataset %>%
  ggplot() +
  geom_histogram(
    mapping = aes(
      x = stress_level,
      y = ..density..),
    bins = 10, 
    fill = "light blue",
    color = "black")+
  labs(
    title = 
      "PMF Distribution of Stress Level")+
  facet_wrap(
    ~heart_attack_risk,
    scales = "free_x")
```


#count of people with heart attack risk among those with no family history.
```{r}
no_family_history_dataset %>%
  group_by(factor(heart_attack_risk)) %>%
  summarize(n = n())
```

#count of people with heart attack risk among those with family history.

```{r}
family_history_dataset %>%
  group_by(
    factor(heart_attack_risk)) %>%
  summarize(n = n())
```

#Barplot showing the distribution of people with different combinations of heart attack risk and family history
#Why is the count messed up!


```{r}
untidied_dataset %>%
  ggplot()+
  geom_bar(
    mapping = aes(
      x = heart_attack_risk)
  )+
  labs(title = "Frequency of Heart Attack Risk")

```

```{r}
untidied_dataset %>%
  ggplot()+
  geom_bar(
    mapping = aes(
      x = family_history)
  )+
  labs(title = "Frequency of Family History")

```
#Modeling (Giselle Rahimi)
#Modeling (exercise_hrs_week)

```{r}
split <- initial_split(no_family_history_dataset, prop = 0.7)
trainData <- training(split)
testData <-testing(split)

```

```{r}
trainData %>%
  summarize(
    total = n(),
    missing = sum(is.na(heart_attack_risk)),
    fraction_missing = sum(is.na(heart_attack_risk)/total)
  )

```

```{r}
model_1 <-
  glm(
  heart_attack_risk ~ exercise_hrs_week,
  data = trainData,
  family = binomial()
)

```

```{r}
model_1_preds <- 
  trainData %>%
  add_predictions(
    model_1,
    type = "response"
  ) %>%
  mutate(
    outcome = 
      if_else(
        pred >= 0.36,
                       true = 1,
                       false = 0))
```

```{r}
model_1_preds %>%
  mutate(
    correct = if_else(
      outcome == heart_attack_risk,
      true = 1,
      false = 0
    )
  ) %>%
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )

```

```{r}
logistic_cv1 <- 
  cv.glm(
    trainData,
    model_1, K =5)

```

```{r}
logistic_cv1$delta
```

#Modeling (sleep_hrs)

```{r}
model_2 <- 
  glm(
  heart_attack_risk ~ sleep_hrs,
  data = trainData,
  family = binomial()
)

```

```{r}
model_2_preds <- 
  trainData %>%
  add_predictions(
    model_2,
    type = "response"
  ) %>%
  mutate(
    outcome = 
      if_else( 
        pred >0.36,
                       true = 1,
                       false = 0))
```

```{r}
model_2_preds %>%
  mutate(
    correct = 
      if_else(
      outcome == heart_attack_risk,
      true = 1,
      false = 0
    )
  ) %>%
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )

```

```{r}
logistic_cv2 <- 
  cv.glm(
    trainData,
    model_2,
    K = 5)

```

```{r}
logistic_cv2$delta
```
#Third model:

```{r}
model_3 <- 
  glm(
  heart_attack_risk ~ stress_level,
  data = trainData,
  family = binomial()
)

```

```{r}
model_3_preds <- 
  trainData %>%
  add_predictions(
    model_3,
    type = "response"
  ) %>%
  mutate(
    outcome = 
      if_else( 
        pred >0.36,
                       true = 1,
                       false = 0))
```

```{r}
model_3_preds %>%
  mutate(
    correct = 
      if_else(
      outcome == heart_attack_risk,
      true = 1,
      false = 0
    )
  ) %>%
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )

```

```{r}

logistic_cv3 <- 
  cv.glm(
    trainData,
    model_3,
    K = 5)

```

```{r}
logistic_cv3$delta
```

#Fourth model:

```{r}
model_4 <- 
  glm(
  heart_attack_risk ~ exercise_hrs_week + sleep_hrs + stress_level,
  data = trainData,
  family = binomial()
)

```

```{r}
model_4_preds <- 
  trainData %>%
  add_predictions(
    model_4,
    type = "response"
  ) %>%
  mutate(
    outcome = 
      if_else(pred >= 0.36,
                       true = 1,
                       false = 0))
```

```{r}
model_4_preds %>%
summarize(
  min = min(outcome),
  max = max(outcome)
)


```

```{r}
model_4_preds %>%
summarize(
  min = min(pred)
)


```

```{r}
model_4_preds %>%
  arrange(desc(pred))
```


```{r}
model_4_preds %>%
  mutate(
    correct = 
      if_else(
      outcome == heart_attack_risk,
      true = 1,
      false = 0
    )
  ) %>%
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )

```

```{r}

logistic_cv4 <- 
  cv.glm(
    trainData,
         model_4,
    K = 5)

```

```{r}
logistic_cv4$delta
```

##Model 5: 

```{r}
model_5 <- 
  glm(
  heart_attack_risk ~ stress_level + exercise_hrs_week,
  data = trainData,
  family = binomial()
)

```

```{r}
model_5_preds <- 
  trainData %>%
  add_predictions(
    model_5,
    type = "response"
  ) %>%
  mutate(
    outcome = 
      if_else( 
        condition = pred >0.36,
                       true = 1,
                       false = 0))
```

```{r}
model_5_preds %>%
  mutate(
    correct = 
      if_else(
      condition = heart_attack_risk == outcome,
      true = 1,
      false = 0
    )
  ) %>%
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )

```


```{r}

logistic_cv5 <- 
  cv.glm(
    trainData,
         model_5,
    K = 5)

```

```{r}
logistic_cv5$delta
```
## Model 1 has the highest accuracy!
## Accuracy and cross validation error on test dataset:

```{r}
model_1_preds_test <- 
  testData %>%
  add_predictions(
    model_1,
    type = "response"
  ) %>%
  mutate(
    outcome = 
      if_else(
        pred >= 0.36,
                       true = 1,
                       false = 0))
```

```{r}
model_1_preds_test %>%
  mutate(
    correct = if_else(
      outcome == heart_attack_risk,
      true = 1,
      false = 0
    )
  ) %>%
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )

```

```{r}
logistic_cv1_test <- 
  cv.glm(
    testData,
    model_1, K =5)

```
```{r}
logistic_cv1_test$delta
```



#Hypothesis testing: (Sukyoung Yoon)

#Change the heart_attack_risk variable to a character object.

```{r}
  no_family_history_dataset$heart_attack_risk <- 
  as.character(
    no_family_history_dataset$heart_attack_risk)

```
#Separated the stress_level variable into two categories (True indicates stress and False indicates no stress)
 
```{r}

no_family_history_dataset <- 
  no_family_history_dataset %>%
  mutate(
    stress =
      ifelse(
        stress_level >= 6,
        "True",
      ifelse(
        stress_level >= 1,
        "False", NA
  )))

```
# Changing the stress variable to a character 

```{r}
  no_family_history_dataset$stress <- 
  as.character(
    no_family_history_dataset$stress)

```

```{r}
str(no_family_history_dataset)
```

#Test for a difference in proportions (True - False) with stress

```{r}
heart_null <- no_family_history_dataset %>%
  specify(
    heart_attack_risk ~ stress,
    success = "1") %>%
  hypothesize(
    null ="independence") %>%
  generate(
    reps = 10000,
    type = "permute") %>%
  calculate(
    stat = "diff in props",
    order = c("True", "False"))

```

```{r}
heart_obs_stat <- 
  no_family_history_dataset %>%
  specify(
    heart_attack_risk ~ stress,
    success = "1") %>%
  calculate(
    stat = "diff in props",
    order = c("True", "False"))
```

```{r}
heart_null %>%
  get_p_value(
    obs_stat = heart_obs_stat,
    direction = "both")
```

```{r}
heart_null %>%
  visualize() +
  shade_p_value(
    obs_stat = heart_obs_stat,
    direction = "both") +
  labs(
    title = "Difference in proportion of stress level",
    x = "Difference in proportion of stress level",
    y = "count")
```


#


```{r}
  no_family_history_dataset <- 
  no_family_history_dataset %>%
  mutate(
    yes_no =
      ifelse(
        heart_attack_risk == 0,
        "yes",
      ifelse(
        heart_attack_risk == 1,
        "no", NA
  )))
  
```

```{r}
unique(
  no_family_history_dataset$sleep_hrs)
```


#separating sleep into high and low

```{r}
no_family_history_dataset <- 
  no_family_history_dataset %>%
  mutate(
    sleep = 
      ifelse(
        sleep_hrs >= 7,
        "High",
      ifelse(
        sleep_hrs < 7,
        "Low", NA
  )))
```


```{r}
heart_null2 <- no_family_history_dataset %>%
  specify(
    yes_no ~ sleep,
    success = "yes") %>%
  hypothesize(
    null = "independence") %>%
  generate(
    reps = 10000,
    type = "permute") %>%
  calculate(
    stat = "diff in props",
    order=c("High", "Low"))
```


```{r}
heart_obs_2 <- 
  no_family_history_dataset %>%
  specify(
    yes_no ~ sleep,
    success = "yes") %>%
  calculate(
    stat = "diff in props",
    order=c("High", "Low"))
```


```{r}
heart_null2 %>%
  get_p_value(
    obs_stat = heart_obs_2,
    direction = "both")
```
```{r}
heart_null2 %>%
  visualize() +
  shade_p_value(
    obs_stat = 
      heart_obs_2,
    direction = "both") +
  labs(
    title = "Difference in proportion of sleep level",
    x = "Difference in proportion of sleep level",
    y = "count")
```

#Hypothesis test difference in proprotions for exercise_hrs:

#Separating exercise hours into high and low categories.

```{r}

no_family_history_dataset <- no_family_history_dataset %>%
  mutate(
    exercise =
      ifelse(
        exercise_hrs_week >= 10,
        "High",
      ifelse(exercise_hrs_week < 10,
             "Low", NA
  )))

```

```{r}
heart_null3 <- 
  no_family_history_dataset %>%
  specify(
    yes_no ~ exercise,
    success = "yes") %>%
  hypothesize(
    null = "independence") %>%
  generate(
    reps = 10000,
    type = "permute") %>%
  calculate(
    stat = "diff in props",
    order=c("High", "Low"))
```


```{r}
heart_obs_3 <- 
  no_family_history_dataset %>%
  specify(
    yes_no ~ exercise,
    success = "yes") %>%
  calculate(
    stat = "diff in props",
    order=c("High", "Low"))
```


```{r}
heart_null3 %>%
  get_p_value(
    obs_stat = heart_obs_3,
    direction = "both")
```
```{r}
heart_null3 %>%
  visualize() +
  shade_p_value(
    obs_stat = heart_obs_3,
    direction = "both") +
  labs(
    title = "Difference in proportion of 
    exercise hour category", 
    x = "Difference in proportion
    of exercise hour category",
    y = "count")
```
